---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
activity = read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
Total Number of steps:
```{r}
sum(activity$steps, na.rm=TRUE)
```

Total Number of steps taken each day:
```{r}
x <- aggregate(activity$steps, by=list(activity$date), FUN=sum, na.rm=TRUE)
print(x)
```

Histogram of the total Steps taken each day
```{r}
hist(x$x, xlab = "Steps", main = "Steps taken each day", col = "azure")
```

Mean  and Median of the total number of steps taken each day:
```{r}
mean(x$x); median(x$x)
```

## What is the average daily activity pattern?
Time Series Plot with the Average number of steps taken by Interval across all dates:
```{r}
t <- aggregate(activity$steps, by=list(activity$interval), FUN=sum, na.rm=TRUE)
plot(t$Group.1, t$x, type = "l", xlab="Interval", ylab="Avg. of Steps per Day")
```

Interval with the Maximum number of steps in average
```{r}
t$Group.1[t$x==max(t$x)]
```

## Imputing missing values
Total number of Missing Values:
```{r}
length(which(is.na(activity$steps)))
```
Rows with missing values are removed from the dataframe:
```{r}
activityWithoutNA <- activity[complete.cases(activity),]
```

Histogram of the total Steps taken each day
```{r}
x2 <- aggregate(activityWithoutNA$steps, by=list(activityWithoutNA$date), FUN=sum)
hist(x2$x, xlab = "Steps", main = "Steps taken each day", col = "azure")
```

Mean  and Median of the total number of steps taken each day:
```{r}
mean(x2$x); median(x2$x)
```

## Are there differences in activity patterns between weekdays and weekends?
Factor Variable that specifies the name of the weekday
```{r}
activityWithoutNA$weekday  <- as.factor(weekdays(strptime(activityWithoutNA$date, "%Y-%m-%d")))
```

time series plots for Weekdays and Weekends
```{r}
par(mfrow = c(2,1))
activityWithoutNAWeekDay <- activityWithoutNA[!(activityWithoutNA$weekday %in% c("Saturday","Sunday")),]
activityWithoutNAWeekEnd <- activityWithoutNA[activityWithoutNA$weekday %in% c("Saturday","Sunday"),]
t2 <- aggregate(activityWithoutNAWeekDay$steps, by=list(activityWithoutNAWeekDay$interval), FUN=sum)
t3 <- aggregate(activityWithoutNAWeekEnd$steps, by=list(activityWithoutNAWeekEnd$interval), FUN=sum)
plot(t2$Group.1, t2$x, type = "l", xlab="Interval", ylab="Number of Steps", main = "Weekdays")
plot(t3$Group.1, t3$x, type = "l", xlab="Interval", ylab="Number of Steps", main = "Weekends")

```